---
title: "Exercise: Panel Data and Fixed Effects"
subtitle: "Session 6: Quantitative Data Analysis: Applied Data Science and Modern Econometrics"
author: "Your name here"
date: "28 May 2026"
format:
  html:
    toc: true
    number-sections: true
    theme: cosmo
    self-contained: true
execute:
  echo: true
  warning: false
  message: false
---

::: {.callout-note}
## Business question
*Does higher workforce education predict better management quality —
and does this relationship hold within firms over time?*

Work through the tasks below. Write your R code in the chunks provided
and your interpretation in the text below each chunk. Write in complete
sentences — imagine you are explaining results to a colleague who knows
regression but is new to panel data.
:::

---

# Setup

Run this chunk first to load packages and data.

```{r}
#| label: setup
library(tidyverse)
library(fixest)
library(modelsummary)
library(flextable)

wms <- read_csv("data/wms_da_textbook.csv") |>
  mutate(log_emp = log(emp_firm))
```

Before starting the tasks, take one minute to orient yourself.

```{r}
#| label: data-orientation
# How many observations, firms, and waves does the panel contain?
# Hint: use n_distinct() and n()
```

**Write here:** Briefly describe what the dataset contains
(units, time dimension, approximate size).

---

# Task 1 — Pooled OLS as a baseline

Estimate a **pooled OLS** model regressing `management` on `degree_nm`
(share of non-managers with a college degree) and `log_emp`
(log number of employees). Use `feols()` with no fixed effects.
Here and in all models below, you should report not only point estimates, but also the 95% confidence interval.

```{r}
#| label: task1-ols
# Your code here
```

**Interpret the coefficient on `degree_nm`:**

*(What does the estimate say? Is it statistically significant?
What is the limitation of this estimate given that it uses pooled data?)*

---

# Task 2 — Introducing firm fixed effects

Now estimate the same model but add **firm fixed effects** using
`| firmid` in `feols()`. Cluster standard errors at the firm level
using `vcov = ~firmid`.

```{r}
#| label: task2-firm-fe
# Your code here
```

**Compare to Task 1:**

*(Did the coefficient on `degree_nm` change? In which direction?
What does this tell you about the role of unobserved firm-level
characteristics — i.e., what was the OVB in Task 1?)*

---

# Task 3 — Two-way fixed effects

Extend the firm FE model by also adding **wave fixed effects**
to control for year-specific shocks. Add `+ wave` to the FE
specification.

```{r}
#| label: task3-twoway-fe
# Your code here
```

**Interpret the change:**

*(Did the coefficient change substantially when adding wave FE?
What kinds of confounders does the wave FE absorb?
Give a concrete example relevant to this dataset.)*

---

# Task 4 — Side-by-side comparison

Use `modelsummary()` to present all three models from Tasks 1–3
in a single table. Report `nobs`, `r.squared`, and `r2.within`
as goodness-of-fit statistics. Add a note about standard errors.

```{r}
#| label: task4-table
# Your code here
# Hint:
# modelsummary(
#   list("Pooled OLS" = m_ols, "Firm FE" = m_fe, "Two-way FE" = m_2way),
#   gof_map = c("nobs", "r.squared", "r2.within"),
#   notes   = "..."
# )
```

**Discuss the within-$R^2$:**

*(What does the within-$R^2$ measure in a fixed effects model?
How does it differ from the overall $R^2$?
Which one should you report, and why?)*

---

# Task 5 — Reflection

::: {.callout-warning}
## Think before you code
This task requires no new code. Read carefully and write 3–5 sentences.
:::

The FE estimator controls for **time-invariant** unobserved heterogeneity.

Think about the variable `ownership` (ownership type: founder, family,
dispersed, etc.).

**Answer the following questions in writing:**

1. What would happen if you tried to include `ownership` as a control
   variable alongside firm fixed effects? Why?

2. Give one example of a **time-varying** confounder that firm fixed
   effects would *not* control for in this dataset. Why is it still
   a potential threat to our interpretation?

---

# Optional Extension A — Restricted sample comparison

::: {.callout-tip}
## Why this matters
When you moved from Task 1 (pooled OLS) to Task 2 (firm FE), two things changed at once:
(1) the estimator, and (2) the sample — from 12,793 to ~5,900 observations, because
singleton firms are dropped. This makes the comparison hard to interpret: is the coefficient
change due to the FE correction, or because multi-wave firms are simply different from
singleton firms?

To find out, run pooled OLS on the **FE-eligible sample only** (firms observed more than once)
and place it between the two previous estimates.
:::

**Step 1:** Create a dataset restricted to firms that appear in at least two waves.

```{r}
#| label: optional-fe-sample
fe_firms <- wms |>
  count(firmid) |>
  filter(n > 1) |>
  pull(firmid)

wms_fe <- wms |> filter(firmid %in% fe_firms)
```

**Step 2:** Run pooled OLS on this restricted sample.

```{r}
#| label: optional-ols-fe-sample
# Your code here
# m_ols_fe <- feols(management ~ degree_nm + log_emp, data = wms_fe)
```

**Step 3:** Build a three-column comparison table: OLS (full) | OLS (FE sample) | Firm FE.

```{r}
#| label: optional-sample-comparison-table
# modelsummary(
#   list("OLS (full)" = m_ols, "OLS (FE sample)" = m_ols_fe, "Firm FE" = m_fe),
#   coef_map  = c("degree_nm" = "Degree (non-managers)", "log_emp" = "Log(employees)"),
#   statistic = "conf.int", conf_level = 0.95,
#   gof_map   = c("nobs", "r.squared", "r2.within"),
#   stars = TRUE
# )
```

**Interpret:** Do the two OLS estimates differ? What does this tell you about whether the
coefficient drop from Task 1 to Task 2 is driven by sample selection or by the FE correction?

*Your answer:*

---

# Optional Extension B — Country fixed effects

::: {.callout-tip}
## If you finish early
Try adding country fixed effects (`| cty`) instead of firm fixed
effects. How does the coefficient on `degree_nm` compare to the
firm FE result? What different question does country FE answer?
Write two sentences explaining the difference in interpretation.
:::

```{r}
#| label: optional-country-fe
# Optional: country FE model
```
